Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-12-29 manuscript/manuscript.tex
Add number-of-tesseroid figure and modify text

Because the density-based discretization algorithm introduces more division along the
radial dimension, it is reasonable that the computation time for variable densities will
be increased in comparison to the homogeneous density case.
In addition to the extra division, there is also the overhead imposed by the extra
function calls to evaluate the density function.
In order to quantify these effects, compare the computation times between the
homogeneous density case and the linear and the exponential density ones.
We note that these results are highly dependent on our particular implementation of the
algorithm and choice of programming language.

We define a $20^\circ \times 20^\circ$ tesseroid of 1 km of thickness and a single
computation point located at different heights above the top surface of the tesseroid (1
m, 10 m, 100 m, 1 km, 10 km, 100 km and 1000 km).
We then compute $V$ and $g_z$ for the homogeneous, linear, and exponential (with $b =
3$) density cases.
We perform 1000 computations for each field, density function, and height while tracking
their computation time.
The final computation time is the mean of these 1000 executions.

Fig.~\ref{fig:computation-times} shows the computation times relative to the homogeneous
case (i.e., ${\Delta t}/{\Delta t_\text{homog}}$).

\begin{figure}
\centering
\iftwocol{
\includegraphics[width=\linewidth]{figures/computation-times.pdf}
}{
\includegraphics[width=0.5\linewidth]{figures/computation-times.pdf}
}
\caption{
    Computation times ratio between linear and exponential density tesseroids with an
    homogeneous one.
}
\label{fig:computation-times}
\end{figure}

>>>>

Because the density-based discretization algorithm introduces more division along the
radial dimension, it is reasonable that the computation time for variable densities will
be increased in comparison to the homogeneous density case.
In order to get a qualitative indicator of this increase in the computation time we
gathered the number of subdivisions the density-based discretization performs under
exponential and sinusoidal density cases.

We assign an exponential density (Eq.~\ref{eq:density-exp}) with values of $b$ equal to
1, 2, 5, 10, 30 and 100, and a sinusoidal density (Eq.~\ref{eq:density-sine}) with
values of $b$ equal to 1, 2, 5, and 10 to an arbitrary tesseroid.
Then we applied the density-based discretization algorithm on each case recording how
many subdivisions it carried out and where the boundaries of the smaller tesseroids are
located.
On both exponential and sinusoidal density cases a value of $\delta$ equal to 0.1 has
been used.

Figs.~\ref{fig:number-of-tesseroids}(a) and~\ref{fig:number-of-tesseroids}(b) show the
exponential and sinusoidal densities used and the resulting boundaries of the
discretized tesseroids as orange dots.
The number of discretizations are independent of the value of factor $b$ in the
exponential density case: the algorithm only performs one split independently of $b$.
On the other hand, the sinusoidal density results show a significant relation between
the number of discretizations and the constant $b$.
Fig.~\ref{fig:number-of-tesseroids} shows an almost linear relation between these two
magnitudes.

Therefore, the computation time for the exponential density would be increased in
comparison to the homogeneous density case because the
single split the density-based algorithm performs, although it will be independent of
the value of $b$.
On the other hand, the computation time for the sinusoidal density is bounded to the
value of $b$ used, greatly increasing it for a large $b$ as 100.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/number-of-tesseroids.pdf}
\caption{
    Number of discretizations performed by the density-based algorithm in case of
    (a)~exponential densities with the same values of $b$ shown in
    Fig.~\ref{fig:exp-densities},
    (b)~sinusoidal densities with the same values of $b$ shown in
    Fig.~\ref{fig:sine-densities}.
    On both figures the boundaries of the discretized tesseroids are marked with orange
    dots.
    Subfigure (c) shows the number of discretized tesseroids for the sinusoidal
    densities as a function of the factor $b$.
}
\label{fig:number-of-tesseroids}
\end{figure*}
